---
title: 'gprofiler: heatmap'
author: "Ronnie Yalung"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(gprofiler2)
library(dplyr)
library(ggplot2)
library(tidyr)
library(plotly)
```

# Data engineering/altering/filtering

```{r}
# assume you already have: protein_data_normalized from "Protein_Data_Pipeline_And_Analysis.rmd"
gpr_data_all <- protein_data_normalized %>%
  dplyr::select(where(is.numeric), GeneSymbol)

# get the names of numeric columns
numeric_cols_gpr <- names(gpr_data_all)[sapply(gpr_data_all, is.numeric)]

# create a list of dataframes removing all 0 values in each column and putting in desc order
gpr_protein_list <- lapply(numeric_cols_gpr, function(colname) {
  gpr_data_all %>%
    dplyr::select(all_of(colname), GeneSymbol) %>%
    filter(.data[[colname]] != 0) %>%
    arrange(dplyr::desc(.data[[colname]]))
})

# name each element in the list by its column
names(gpr_protein_list) <- numeric_cols_gpr
```

# Function for gost results
```{r}
run_gost_from_input <- function(g_protein_list) {
  # Step 1: show available protein names
  cat("Available proteins:\n")
  print(names(g_protein_list))
  
  # Step 2: ask for user input
  protein_input <- readline(prompt = "\nEnter ONE sample name from the list above: ")
  
  # Step 3: validate protein name
  if (!(protein_input %in% names(g_protein_list))) {
    stop(paste0("Error: '", protein_input, "' is not a valid protein name.\n",
                "Valid options are: ", paste(names(g_protein_list), collapse = ", ")))
  }
  
  # Step 4: ask for source input (comma-separated)
  cat("\nAvailable sources include: GO:BP, GO:MF, GO:CC, KEGG, REAC, WP, CORUM, TF, MIRNA, HPA, HP\n")
  source_input <- readline(prompt = "Enter one or more sources (comma-separated): ")
  
  # Parse the source input into a character vector
  sources <- trimws(strsplit(source_input, ",")[[1]])
  
  # Step 5: validate at least one source, CAN CHANGE SOURCES HERE
  valid_sources <- c("GO:BP", "GO:MF", "GO:CC", "KEGG", "REAC", "WP", "CORUM", "TF", "MIRNA", "HPA", "HP")
  if (!all(sources %in% valid_sources)) {
    invalid <- sources[!sources %in% valid_sources]
    stop(paste0("Invalid source(s): ", paste(invalid, collapse = ", "),
                "\nValid options are: ", paste(valid_sources, collapse = ", ")))
  }
  
  # Step 6: retrieve the gene symbols for that protein
  gpr_filter_df <- g_protein_list[[protein_input]]
  gpr_list <- list(gpr_filter_df$GeneSymbol)
  
  # Step 7: run gprofiler2::gost
  # --- Nested safe_gost function (retry logic) ---
  safe_gost <- function(query, sources, max_attempts = 3) {
    for (i in 1:max_attempts) {
      result <- try(
        gost(
          query = query,
          organism = "hsapiens",
          ordered_query = TRUE,
          significant = TRUE,
          sources = sources
        ),
        silent = TRUE
      )
      if (!inherits(result, "try-error") && !is.null(result)) {
        return(result)
      }
      cat("Attempt", i, "failed — retrying in 2 seconds...\n")
      Sys.sleep(2)
    }
    stop("g:Profiler request failed after multiple attempts (API doesn't always work. Maybe retry.)")
  }
  
  # Step 7: run gost safely
  gost_results <- safe_gost(gpr_list, sources)
  
  # Step 8: return results
  cat("\n✅ g:Profiler analysis complete!\n")
  return(gost_results)
}
```

# RUN FUNCTION
```{r}
# get gost results using function
gost_results <- run_gost_from_input(gpr_protein_list)
```



# Data analysis

# GOST PLOT
```{r}
# Run for gost plot
gostplot(gost_results, capped=TRUE, interactive=TRUE)
```

# HEATMAP
```{r}
# Run (this cell and onwards) for heatmap  (term_name and p-value)

# Extract result table
g_res <- gost_results$result
```

# Functions to run on the data (top N significant terms and -log10 the p values)
```{r}
# Optional, simplify p-values: use -log10 transform for visualization (0.1 = 1, 0.01 = 2, 0.001 = 3, etc)
g_res <- g_res %>%
  mutate(neg_log_p = -log10(p_value))

# Optionally, take top N most significant terms
top_terms <- g_res %>%
  arrange(p_value) %>%
  slice_head(n = 30) #TODO: make it so where its not just top 30 overall, but top 30 PER set (e.g. top 30 for go;bp, top 30 for kegg, etc)
```

```{r}
# replace g_res with top_terms, replace p_value with neg_log_p
ggplot(top_terms, aes(x = source, y = reorder(term_name, neg_log_p), fill = neg_log_p)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(
    title = "Top Enriched Terms",
    x = "Source",
    y = "Term Name",
    fill = "p-value" #-log10(p-value)
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    plot.title = element_text(hjust = 0.5)
  )
```
## GOSemSim
# TODO: flip it around. we want multiple SAMPLES and one ontology. note: the HCF_S etc are SAMPLES, not proteins. same for heatmap and gost results. (comparing samples... we want to see the GO:BP for HCF_S sample, then sample 2, and compare those)

# then, do gene concept network plots. then gosemsim.