#-------------------------------------------------------------------------------------------------------------------
#-------------------------------------------------------------------------------------------------------------------
#-------------------------------------------------------------------------------------------------------------------
# This section will load the .xlsx file data, and then create the dataframes necessary for the subsequent functions
# Will load into a large data variable named "data". `data` will have all of the different types of data formats for the functions

## First, load libraries
```{r}
source("pre_interface_functions/data_load_and_transform/all_func.R")
source("pre_interface_functions/protein_enrichment/all_func.R")
source("pre_interface_functions/filtering/all_func.R")
source("pre_interface_functions/plots/all_func.R")
```

## Load .xlsx file into `data` variable. Only .xlsx supported at the moment
```{r}
# Load data into `data` variable
data <- list()
data <- load_dataset(data, "raw_sample_data") # load_data.R; works with .xlsx files
#data <- load_dataset(data, "DATA2")
#view(data$raw_sample_data)
```


## Clean data (Note: data needs to be formatted like the example .xlsx image)
- The `clean_data` function only works with the example .xlsx. 
- You don't have to run either of these functions if the data is already ready to go.
```{r}
# clean .xlsx
data[["clean_data"]] <- clean_data(data$raw_sample_data) # initial_clean.R
```


## Add columns for other identifiers and convert UniProtIds to all other relevant biological identifiers
## -> Some Ids won't match (using bitr from clusterProfiler). Separate those.
Skip this if you don't want the other identifier names, or if you already have them, or if you don't have a uniprotID column
```{r}
data[["clean_data"]]$analysis_data <- convert_uniprot_ids(data$clean_data$cleaned_data) # convert_uniprot_ids.R
data$clean_data <- split_unmapped_ids(data$clean_data, cleaned_name = "analysis_data")
```


# normalize data (choose type, and column numbers)
Can choose the log base (3rd argument) and the numeric columns (2nd argument)
```{r}
data[["analysis_data"]] <- normalize_log(data$clean_data$analysis_data, 2:11, 2)
```


# Run this, if you don't want to normalize the data. Puts analysis_data in a nicer location
```{r}
#data[["analysis_data"]] <- data$clean_data$cleaned_data

# Change names so they work in other functions

#names(data$DATA2)[names(data$DATA2) == "UNIPROT ID"] <- "UniProtID"
#names(data$DATA2)[names(data$DATA2) == "Gene Name"] <- "GeneSymbols"
```


## GO annotation retrieval (GO:Ids, GO names, ontology)
```{r}
# Requires UniProtIDs. function as follows=> add_go_annotation(df, uniprot_colname)
data[["analysis_data"]] <- add_go_annotations(data[["analysis_data"]], uniprot_col = "UniProtID")
data[["DATA2"]] <- add_go_annotations(data[["DATA2"]], uniprot_col = "UniProtID")
```



## Create Numeric DF(s)
These numeric dataframes are required for the structure of below functions and analysis
```{r}
# Can make any numeric DF. Replace 'gene_symbols' with any new df you want created.
# Can choose multiple columns by c("GeneSymbol", "EntrezID"), etc. in the second parameter
data[["numeric_data"]]$analysis_gene_symbols <- create_numeric_df(data$analysis_data, "GeneSymbol")
data[["numeric_data"]]$DATA2_gene_symbols <- create_numeric_df(data$DATA2, "GeneSymbol")


#data[["numeric_data"]]$all_identifiers <- create_numeric_df(data$analysis_data, c("GeneSymbol", "EntrezID", "UniProtID"))
```
#-------------------------------------------------------------------------------------------------------------------
#-------------------------------------------------------------------------------------------------------------------
#-------------------------------------------------------------------------------------------------------------------







# Now, we have the data cleaned, normalized, and properly transformed and structured to include all identifiers for (valid) proteins in our list.
# From here, we can do filtering, enrichment, and plotting, in any order.

## Gprofiler: GOST enrichment (numeric gene symbol)
```{r}
# Can add many gost results. replace 'all_sample_go_bp' with any name you want

# NOTE: Filtering Top N does it BEFORE enrichment analysis

# CRITICAL: this particular GOST requires genesymbol numeric_data

data[["gost_results"]]$data1_all_samples_go_bp <- run_gost_menu(data$numeric_data$analysis_gene_symbols)
data[["gost_results"]]$data2_all_samples_go_bp <- run_gost_menu(data$numeric_data$DATA2_gene_symbols)
```






## Xref with Gprofiler Gost Results (analysis data and gost results)
This basically just cross references our original analysis data with our gost results, and keeps only the data that were in both sets (i.e. analysis_data that is significant). Checks via GO:ID

Below are examples of use cases. use for reference, as the arguments/parameters for the functions can get long and convoluted
```{r}
# one sample
data[["xref_gost"]]$hnpcc_cl <- filter_proteins_by_gost(data$analysis_data, data$gost_results$all_samples_go_bp$HNPCC_CL, sample_col = "HNPCC_CL")
```

```{r}
# all samples
data[["xref_gost"]]$data1_all_samples <- filter_all_sample_by_go(data$analysis_data, data$gost_results$data1_all_samples_go_bp)
data[["xref_gost"]]$data2_all_samples <- filter_all_sample_by_go(data$DATA2, data$gost_results$data2_all_samples_go_bp)
```

```{r}
# number of top proteins you want;... sort descending
data[["xref_gost"]]$top20_hnpcc_cl <- head(data$xref_gost$hnpcc_cl$filtered_data[order(data$xref_gost$hnpcc_cl$filtered_data$HNPCC_CL, decreasing = TRUE), ], 20)
```

```{r}
# filter by expression value cutoff (note, these values may be normalized from their original values (log2))
data[["xref_gost"]]$expr_above_5_hnpcc_cl <- data$xref_gost$hnpcc_cl$filtered_data[data$xref_gost$hnpcc_cl$filtered_data$HNPCC_CL >= 5, ]
```



## Numeric DFs after xref (xref data)
```{r}
data[["numeric_data"]]$data2_test <- create_numeric_df(data$xref_gost$data2_all_samples$Supermere_2D_S7$filtered_data, "GeneSymbol")

# Generate the new data, keep renaming the set.
#Supermere_2D_S8 <- create_numeric_df(data$xref_gost$data2_all_samples$Supermere_2D_S8$filtered_data, "GeneSymbol")

# Append the new list to the existing one
#data[["numeric_data"]]$data2_test <- c(data[["numeric_data"]]$data2_test, Supermere_2D_S8)

# Process all 14 samples in data2_all_samples automatically
# 1. Map the function across all samples to get a list of lists
temp_list <- base::lapply(data$xref_gost$data2_all_samples, function(sample) {
  create_numeric_df(sample$filtered_data, "GeneSymbol")
})

# 2. Use do.call to 'c' (combine) them all into one flat list
# This mimics: c(sample7_list, sample8_list, sample9_list...)
data[["numeric_data"]]$data2_test <- base::do.call(base::c, temp_list)
```




## Gprofiler: Enrichment plots (Gostplot, Heatmap) (gost results)
```{r}
# GOST PLOT --> I believe displayed significance in terms of p-value
view_gost_plots_menu(data$gost_results$all_samples_go_bp)
# or like: view_gost_plots_menu(data$gost_results$all_samples_go_bp$HNPCC_CL)

# HEATMAP

# Filter Top N per group options:
# y = shows the top n genes per sample.
# n = shows the top n genes globally. so if the top 8 are only in HCF_S, it'll show just that. if part of that top 8 is in more than one sample, it'll show just those samples. (not useful in our case? not sure, but it's similar to what I had originally so I decided to keep it as a functionality)
plot_heatmap_gost_menu(data$gost_results$all_samples_go_bp)
# or like: plot_heatmap_gost_menu(data$gost_results$all_samples_go_bp$HNPCC_CL)
```




## Go Semantic Similarity Analysis (numeric dfs)
```{r}
# shouldn't be any data loss since we got the symbols from the same database

# need the $numeric_data$all_identifiers variable, so then we ensure the max amount of coverage (sometimes, a database won't have data for a particular identifier)

data[["semantic_similarity"]]$data2_one <- run_sample_semantic_analysis(data$numeric_data$data2_test)

DOSE::simplot(data[["semantic_similarity"]]$data2_comb)
DOSE::simplot(data[["semantic_similarity"]]$data2_one)
```





## Gene Concept Network with GSEA (numeric data)
NOTE: ISN'T CURRENTLY WORKING BECAUSE OF CHANGES IN THE LIBRARY (enrichplot package v1.28+ have been changed significantly)
```{r}
#get proper data layout
data[["gene_concept_network"]]$all_samples_data <- create_gcn_data(data$numeric_data$all_identifiers)

# run for plots: shouldn't be any data loss since we got the symbols from the same database
data[["gene_concept_network"]]$plots_hnpcc_ex_go_bp <- run_gsea_cnet(data$gene_concept_network$all_samples_data$HNPCC_EX)

# display plots
data$gene_concept_network$plots_hnpcc_ex_go_bp$p5
```
